### 2019-06-13 thesis-template/msc-thesis.tex
update

\section{Objectives}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/hyperstreamlines.png}
  }
  \caption{Hyperstreamlines in tensor field visualization}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{http://www2.cs.uh.edu/~chengu/Teaching/Fall2018/Lecs/Lec17.pdf}}	
  \label{Skalierung}
\end{figure}
The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies, e.g., clustering, structuring or projection techniques. When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor-medical resonance imaging) when measuring water molecule diffusion characteristics within tissue or in simulated (or measured) stress tensor fields from solid and fluid mechanics. The problem setting encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids consisting of mainly second-order tensors (cf. Cauchy stress tensor). The objective of this work is to design a simple light propagation scheme yet following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation with the possibility to model light transport transport in a geometrical scene. This propagation scheme can either be used to visualize the light transport paths (defined by separatrices) in a scene defined by geometry. Or, on the other hand, be used to visualize the anisotropy characteristics of a given tensor field. However, the tensor field needs to be pre-processed somehow to generate a tensor-induced footprint map of transmission profiles. Hence, the tensors need to be decomposed to obtain a kind of transmission profile. Another aim is to implement an approach borrowed from vector-field visualization. We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian coherent structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of varying directional distributions (w.r.t. direction/location) represented by tensors. The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions and since we effectively measure the finite-time Lyapunov exponent (FTLE) with both forward and backward integration time (stimulus directions). The real part of the eigenvalues could be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the symmetric stress tensor in equilibrium, just the principal normal stresses. Thus, we can even define the full set of stresses with the singular values which are derived from singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order). The main goal of this work is therefore the encoding of a tensor field into certain elaborate representations and the interpretation of meaningful results comparing the proposed techniques against others. The designed approach could e.g. be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations.

@@

The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig. \ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$ arranged in row-major order. It appears in similar form as viscous stress tensor in fluid mechanics:
\begin{align*}
\boldsymbol{\sigma}
 = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
\boldsymbol{T}^{(\mathbf{e}_2)} \\
\boldsymbol{T}^{(\mathbf{e}_3)} \\
\end{matrix}}\right] =
\left[{\begin{matrix}
\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
\end{matrix}}\right]  = \left[{\begin{matrix}
\sigma _x & \tau _{xy} & \tau _{xz} \\
\tau _{yx} & \sigma _y & \tau _{yz} \\
\tau _{zx} & \tau _{zy} & \sigma _z \\
\end{matrix}}\right] =
\begin{bmatrix}
\sigma _{11} & \sigma _{12} & \sigma _{13} \\
\sigma _{21} & \sigma _{22} & \sigma _{23} \\
\sigma _{31} & \sigma _{32} & \sigma _{33} 
\end{bmatrix}.
\end{align*}
These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in direction ${x,y,z}$ ($3^2=9$ individual numbers). Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space. The stress tensor itself can be interpreted as a transformation which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$. The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent (kind of a Euclidean distance of the stress components). The tensor itself is a complicated representation and needs to be projected onto its principal axes to lower its dimensionality. It is possible to find 2 principal stress directions in 2D through PCA (see sect. \ref{sec:pca} below), for which there are no shear stresses existent in equilibrium. These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour. The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''. So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig. \ref{HAT} in sect. \ref{sec:pca}.
Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle. It is therefore somehow representing the distortion of an absolutely isotropic object (unit circle) by whatever the affine tensor transformation effects (e.g. stretching, shearing, mirroring, scaling, rotating). Let us consider it as some lower dimensional, more intuitive representation of the tensor which we were left with in the first place.
\section{Principal Component Analysis}
\label{sec:pca}
Note that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following. Also keep in mind, that we interpret a matrix or tensor as affine transformation. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by, e.g., ellipsoid glyphs.

\begin{figure}[!t]
\hskip 22pt
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/HAT.png}
    \label{HAT}
   \caption{2D Principal component ellipsoid (eigenframe)}
   \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:HAT-ellipse0.svg}}
  \end{minipage} \hskip 40pt
  %\raggedright
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/glyphs.png}
    \caption{3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
    \label{3Dglyphs}
    \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://slideplayer.com/slide/5291764/}}
  \end{minipage}
\label{gradient}
\end{figure}

Principal Component Analysis is an algorithm which can capture the principal components (directions) in stochastic data. But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling. Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices. For (skew-) symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf. \cite{kieburg}). Note, that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases. Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of transformation matrix as pointed out by Moler \cite{moler}. The singular value decomposition is defined as follows:
\begin{align*}
	\mathbf{A} = \mathbf{U} \boldsymbol{\Sigma} \mathbf{V}
\end{align*}
where
\begin{itemize}
	\item $\mathbf{U}$ is an $m × m$ unitary matrix
	\item $\boldsymbol{\Sigma}$ is a diagonal matrix {$m × n$} with non-negative real numbers on the diagonal
	\item $\mathbf{V}$ is an {$n × n$} [[unitary matrix]], $\mathbf{V}^*$ is conjugate transpose
\end{itemize}

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/rings.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/tensorfieldlines.png}
    \label{b)}
  \end{minipage}
\caption{a) ``rings'' test field, b) tensor field lines for a)}
\label{rings}
\end{figure}
The ``rings'' test field, depicted in Fig. \ref{rings} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/spiral.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/spiral-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``spiral''-testfield, b) tensor field lines for a)}
\label{spiral}
\end{figure}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/inverse.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/inverse-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``inverse''-testfield, b) tensor field lines for a)}
\label{inverse}
\end{figure}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/corner.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/corner-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``corner''-testfield, b) tensor field lines for a)}
\label{corner}
\end{figure}

@@

There has been extensive recent work on symmetric tensors but in comparison little on asymmetric ones. Since we aren't able to cover the whole scope in this work, we will focus on the most relevant works on tensor field visualization. We will talk about the publications from Hlawatsch et al. \cite{hlawatsch} and Zirr et al. \cite{zirr} in particular, since they pose thematically related works.\vskip 5pt

@@

\section{Requirements and Ambitions}
\label{sec:spec} 
Our ambition is to motivate a simple and efficient global illumination propagation scheme able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t. energy conservation and propagation attenuation principles. We then generate transmission profiles from the eigensystems (principal component frames) of the tensors for each cell in the domain (field). That is, to provide an orientation for a grid of crystal fibre structures (cf. optical fibres) to redirect the intensities in analogy to the anisotropies of the tensor field within a glyph representation, obtained from Principal Component Analysis. Our approach should be implicitly adaptive to any kind of 2D dataset. Thus, we require to be compatible to handle any kind of matrix (symm./anti-symm.) and any kind of resolution of the field (full-spectrum). At last, our ambition is to segment key locations (LCS/ridges) in the field with tensor field lines (hyperstreamlines) converging/diverging, which is necessarily the same (for bidirectional eigenvectors of tensors). Thus, we aim to generate a new FTLE-related method for tensor field visualization by applying a light transport gradient to the resulting light distributions for every possible position and direction of the light source in the grid generated by stimulation with Delta-pulses. Our approach should efficiently handle (relatively) large datasets with reasonable resolutions. 
%Knapp zwei Seiten, in dem die Anforderungen, die Zielsetzung und die
%Methoden überblicksartig beschrieben werden. Hier sollte die
%Beschreibung ``technischer'' bzw.~``formaler'' sein als in der
%Einleitung, da der Leser nun mit den Grundlagen und verwandten
%Arbeiten vertraut ist.

\section{Light Transport (Propagation) Scheme}
\label{sec:scheme}
Within the scope of this work, we will motivate a light transport propagation scheme designed to efficiently propagate the initial intensity profiles stored in polar coordinates given as symbolic impulses (stimuli) by the user or the config. Polar profiles for different light sources are shown in Fig. \ref{polar}:
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.8\textwidth]
    {img/strahler.png}
  }
  \caption{Polar profiles for different types of emitters ($m=0$: spherical emitter, $m=1$: Lambert emitter, $m=2$: cone emitter}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{http://www.optik.uni-erlangen.de/odem/vorlesung/ws1213/Photometrie_Geometrische_Optik.pdf}}
  \label{polar}
\end{figure}\vskip 3pt
As a reminder, polar profiles map each angle $\varphi$ a magnitude $r$ forming a function $r(\varphi)$. They are circular axis versions of the domain $[0..2\pi]$, with negative magnitudes mirrored by 180 degrees). Thus, they are functions with an angular domain making them directionally dependent, which is a well suited representation for point lights, as we use them in a style of Dachsbacher et al. \cite{dachsbacher}. For the sake of interpretability, we exclude negative values (energies) from the scope of our calculations which we express through mathematical clip functions, limiting negative values to $0.0$, which we denote by subscript $+$ in the following. As a start-up thinker for the propagation scheme, we will think of a grid cell in 2D as a center pixel within an 8-neighborhood. Remember, that all considerations from now on are given concerning a single grid cell respectively. It has 8 (4 faces/edges and 4 diagonals) unique neighbors, which are adjacent point lights in the grid (the type of light source with theoretically no finite extend). We consider one face neighbor (top) to shortly explain the steps of the scheme. The diagonals and the others are obtained from derivation and symmetry.

Each cone (yellow and green) is an individual neighbor-dependent angular area. We optimize for distributing contributions for overlapping cones to transmit attenuated (because of the larger factor $\sqrt{2}$ distance) intensity in form of a shared part to the diagonals in a single step (which prevents from zigzagging diagonal intensities). Therefore, the intensity gathered in polar coordinates in areas marked in green, needs to be partitioned between 2 of the 8 cone-index neighbors with following considerations:

@@

We place a cosine lobe scaled with the integrated (accumulated) energies in each of the 8 cone directions, whereas outgoing integrated and summed energies are weighted by the linear combination weights (factors) in overlapping (green) areas to split up the contributions accordingly. The contribution areas are depicted in Fig. \ref{scheme}, the cosine lobes can be observed in Fig. \ref{iter} respectively. This manner of propagation is inspired by the propagation scheme of Dachsbacher et al. \cite{dachsbacher} (see the paper on how to ``project the flux into a point light''). The implementation is done in a dual buffer approach which pushes the energies back and forth (buffer A to buffer B) until convergence, which happens when the energy is spreaded throughout the grid and enters a stationary state, characterized by equal out (at grid borders) to in (at light source positions) energy-flow with no more nettings going on inside the domain. We measure the convergence criterion by setting up an overall distribution error w.r.t to the previous iteration. Note that even though we propagate the energy directly through the diagonal cones (which forms a square initially), the 8 degrees of freedom introduce a circular (spherical) wave front after few iterations (for greater resolutions) as depicted in Fig. \ref{iter}. This occurs, since the diagonals are more strongly attenuated by default and there is a redistribution of intensities in the long term resulting in a circular torus-shaped distribution, as also observed in Fig. \ref{iter}. Also note, that this approach satisfies energy conservation and propagation attenuation principles and applies to the principles of light propagation in vacuum (by default). The verification of these principal requirements and specifications can be observed in the evaluation in chap. \ref{chap:eval}.

>>>>

\section{Objectives}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/hyperstreamlines.png}
  }
  \caption{Hyperstreamlines in tensor field visualization}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{http://www2.cs.uh.edu/~chengu/Teaching/Fall2018/Lecs/Lec17.pdf}}	
  \label{Skalierung}
\end{figure}
The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies, e.g., clustering, structuring or projection techniques. When it comes to tensor fields, this comes in handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor-medical resonance imaging) when measuring water molecule diffusion characteristics within tissue or in simulated (or measured) stress tensor fields from solid and fluid mechanics. In this  domains of science the problem arises, that tensor fields can be generated (simulated) or measured. The problem setting encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids consisting of mainly second-order tensors (cf. Cauchy stress tensor). The objective of this work is to design a simple light propagation scheme yet following the most basic but crucial physical principles and laws, like energy conservation and propagation attenuation with the possibility to model light transport transport in a geometrical scene. This propagation scheme can either be used to visualize the light transport paths (defined by separatrices) in a scene defined by geometry. Or, on the other hand, be used to visualize the anisotropy characteristics of a given tensor field. However, the tensor field needs to be pre-processed somehow to generate a tensor-induced footprint map of transmission profiles. Hence, the tensors need to be decomposed to obtain a kind of transmission profile. Another aim is to implement an approach borrowed from vector-field visualization. We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian coherent structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of varying directional distributions (w.r.t. direction/location) represented by tensors. The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions and since we effectively measure the finite-time Lyapunov exponent (FTLE) with both forward and backward integration time (stimulus directions). The real part of the eigenvalues could be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the symmetric stress tensor in equilibrium, just the principal normal stresses. Thus, we can even define the full set of stresses with the singular values which are derived from singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order). The main goal of this work is therefore the encoding of a tensor field into certain elaborate representations and the interpretation of meaningful results comparing the proposed techniques against others. The designed approach could e.g. be applied in medicinal sciences, engineering and physics to reveal structures which are hard to grasp through interpreting simpler and sometimes confusing glyph and tensor field line visualizations.

@@

The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig. \ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$ arranged in row-major order. It appears in similar form as viscous stress tensor in fluid mechanics:
\begin{align*}
\boldsymbol{\sigma}
 = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
\boldsymbol{T}^{(\mathbf{e}_2)} \\
\boldsymbol{T}^{(\mathbf{e}_3)} \\
\end{matrix}}\right] =
\left[{\begin{matrix}
\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
\end{matrix}}\right]  = \left[{\begin{matrix}
\sigma _x & \tau _{xy} & \tau _{xz} \\
\tau _{yx} & \sigma _y & \tau _{yz} \\
\tau _{zx} & \tau _{zy} & \sigma _z \\
\end{matrix}}\right] =
\begin{bmatrix}
\sigma _{11} & \sigma _{12} & \sigma _{13} \\
\sigma _{21} & \sigma _{22} & \sigma _{23} \\
\sigma _{31} & \sigma _{32} & \sigma _{33} 
\end{bmatrix}.
\end{align*}
These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in direction ${x,y,z}$ ($3^2=9$ individual numbers). Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space. The stress tensor itself can be interpreted as a transformation which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$. The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent (kind of a Euclidean distance of the stress components). The tensor itself is a complicated representation and needs to be projected onto its principal axes to lower its dimensionality. It is possible to find 2 principal stress directions in 2D through PCA (see sect. \ref{sec:pca} below), for which there are no shear stresses existent in equilibrium. These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour. The resulting system of principal directions and absolute stresses is often called an ``eigensystem'' or ``eigenframe''. So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig. \ref{HAT} in sect. \ref{sec:pca}.
Another interpretation for the principal component ellipsoid is that it is the result of the tensor transformation applied to the unit circle. It is therefore somehow representing the distortion of an absolutely isotropic object (unit circle) by whatever the affine tensor transformation effects (e.g. stretching, shearing, mirroring, scaling, rotating). Let us consider it as some lower dimensional, more intuitive representation of the tensor which we were left with in the first place. To put it all together, a tensor is a mathematical transformation which is encoded inside of an abstract $oD$ number array, which comes in need, whenever we have a whole distribution of directions manifesting itself in a transformation ellipsoid.

\section{Polar Coordinates}
\label{sec:polar}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.7\textwidth]
    {img/polarkoordinaten.png}
  }
  \caption{polar space}
  \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:Ebene_polarkoordinaten.PNG}}
  \label{polar}
\end{figure}\vskip 3pt
Mathematically speaking, a polar space is a system which maps each angle $\omega$ a magnitude $r(\omega)$ through a symbolic function. The conversion formula from polar (given a radius $r$ and angle $\omega$) to Cartesian coordinates is depicted in Fig. \ref{polar} and is given by:
\begin{align*}
	x &= r \cos \varphi, \\
 	y &= r \sin \varphi.
\end{align*}
The inverse direction (given coordinates $x,y$) is a bit more complicated:
\begin{align*}
	r &= \sqrt{x^2 + y^2} \\
\varphi &= \operatorname{atan2}(y, x)
\end{align*}
The polar domain is a circular axis version of the domain $[0..2\pi]$, with negative magnitudes shifted by 180 degrees (point reflected) as this is the formal behaviour defined by the conversion formulas.

\section{Principal Component Analysis}
\label{sec:pca}
Note that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following. Also keep in mind, that we interpret a matrix or tensor as affine transformation. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by, e.g., ellipsoid glyphs.

\begin{figure}[!t]
\hskip 22pt
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/HAT.png}
    \label{HAT}
   \caption{2D principal component ellipsoid (eigenframe)}
   \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://de.wikipedia.org/wiki/Datei:HAT-ellipse0.svg}}
  \end{minipage} \hskip 40pt
  %\raggedright
  \begin{minipage}[!t]{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/glyphs.png}
    \caption{3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
    \label{3Dglyphs}
    \captionsetup{font={footnotesize,bf,it}}
   \caption*{src: \url{https://slideplayer.com/slide/5291764/}}
  \end{minipage}
\label{gradient}
\end{figure}

Principal Component Analysis is an algorithm which can capture the principal components (directions) in stochastic data. But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling. Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices. For (skew-) symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf. \cite{kieburg}). Note, that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases. Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of transformation matrix as pointed out by Moler \cite{moler}. The singular value decomposition is defined as follows:
\begin{align*}
	\mathbf{A} = \mathbf{U} \boldsymbol{\Sigma} \mathbf{V}
\end{align*}
where
\begin{itemize}
	\item $\mathbf{U}$ is an $m × m$ unitary matrix
	\item $\boldsymbol{\Sigma}$ is a diagonal matrix {$m × n$} with non-negative real numbers on the diagonal
	\item $\mathbf{V}$ is an {$n × n$} unitary matrix, $\mathbf{V}^*$ is conjugate transpose
\end{itemize}

@@

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/rings.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/tensorfieldlines.png}
    \label{b)}
  \end{minipage}
\caption{a) ``rings'' test field, b) tensor field lines for a)}
\label{rings}
\end{figure}
The ``rings'' test field, depicted in Fig. \ref{rings} should be used to demonstrate the directive ability of the approach transmitting energy in circular orbits.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/spiral.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/spiral-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``spiral''-testfield, b) tensor field lines for a)}
\label{spiral}
\end{figure}
The ``spiral'' test field, depicted in Fig. \ref{spiral} should be used to demonstrate the directive ability of the approach transmitting energy in archimedian spiral orbits.
\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/inverse.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/inverse-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``inverse''-testfield, b) tensor field lines for a)}
\label{inverse}
\end{figure}
With this test field we want to demonstrate the directive separation of intensities into four quadrants.

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/corner.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=\textwidth]{img/corner-TFL.png}
    \label{b)}
  \end{minipage}
\caption{a) ``corner''-testfield, b) tensor field lines for a)}
\label{corner}
\end{figure}
The test field ``corner'' poses a sustained turn over the grid representing the redirection or forwarding along domain corners.

@@

There has been extensive recent work on symmetric tensors but in comparison little on asymmetric ones. Since we aren't able to cover the whole scope in this work, we will focus on the most relevant works on tensor field visualization. We will talk about the publications from Hlawatsch et al. \cite{hlawatsch} and Zirr et al. \cite{zirr} in particular, since they pose thematically close related works.\vskip 5pt

@@

\section{Requirements and Ambitions}
\label{sec:spec} 
Our ambition is to motivate a simple and efficient global illumination propagation scheme able to distribute energy profiles in discrete polar form in 2D/3D-space w.r.t. energy conservation and propagation attenuation principles. We then generate transmission profiles from the eigensystems (principal component frames) of the tensors for each cell in the domain (field). This step is done, to provide an orientation for a grid of crystal fibre structures (cf. optical fibres) to redirect the intensities in analogy to the anisotropies of the tensor field within a glyph representation, obtained from Principal Component Analysis. Our approach should be implicitly adaptive to any kind of 2D dataset. Thus, we require to be compatible to handle any kind of matrix (symm./anti-symm.) and any kind of resolution of the field (full-spectrum). At last, our ambition is to segment key locations (LCS/ridges) in the field with tensor field lines (hyperstreamlines) converging/diverging, which is necessarily the same (for bidirectional eigenvectors of tensors). Thus, we aim to generate a new FTLE-related method for tensor field visualization by applying a light transport gradient to the resulting light distributions for every possible position and direction of the light source in the grid generated by stimulation with Delta-pulses. Our approach should efficiently handle (relatively) large datasets with reasonable resolutions. 
%Knapp zwei Seiten, in dem die Anforderungen, die Zielsetzung und die
%Methoden überblicksartig beschrieben werden. Hier sollte die
%Beschreibung ``technischer'' bzw.~``formaler'' sein als in der
%Einleitung, da der Leser nun mit den Grundlagen und verwandten
%Arbeiten vertraut ist.

\section{Light Transport (Propagation) Scheme}
\label{sec:scheme}
Within the scope of this work, we will motivate a light transport propagation scheme designed to efficiently propagate the initial intensity profiles stored in polar coordinates given as symbolic impulses (stimuli) by the user or the config. Polar profiles for different light sources are shown in Fig. \ref{polar}:
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.8\textwidth]
    {img/strahler.png}
  }
  \caption{Polar profiles for different types of emitters ($m=0$: spherical emitter, $m=1$: Lambert emitter, $m=2$: cone emitter}
  \captionsetup{font={footnotesize,bf,it}}
  \caption*{src: \url{http://www.optik.uni-erlangen.de/odem/vorlesung/ws1213/Photometrie_Geometrische_Optik.pdf}}
  \label{polar}
\end{figure}\vskip 3pt
As a reminder, polar profiles map each angle $\varphi$ a magnitude $r$ forming a function $r(\varphi)$. They are circular axis versions of the domain $[0..2\pi]$, with negative magnitudes mirrored by 180 degrees. Thus, they are functions with an angular domain making them directionally dependent, which is a well suited representation for point lights, as we use them in a style of Dachsbacher et al. \cite{dachsbacher}. For the sake of interpretability, we exclude negative values (energies) from the scope of our calculations which we express through mathematical clip functions, limiting negative values to $0.0$, which we denote by subscript $+$ in the following. As a start-up thinker for the propagation scheme, we will think of a grid cell in 2D as a center pixel within an 8-neighborhood. Remember, that all considerations from now on are given concerning a single grid cell respectively. It has 8 (4 faces/edges and 4 diagonals) unique neighbors, which are adjacent point lights in the grid (the type of light source with theoretically no finite extend). We consider one face neighbor (top) to shortly explain the steps of the scheme. The diagonals and the others are obtained from derivation and symmetry.

Each cone (yellow and green) is an individual neighbor-dependent angular area. We optimize for distributing contributions for overlapping cones to transmit attenuated (because of the larger factor $\sqrt{2}$ distance) intensity in form of a shared part to the diagonals in a single step (which prevents from zigzagging diagonal intensities). Therefore, the intensity gathered in polar coordinates in areas marked in green, needs to be partitioned between 2 of the 8 cone index neighbors with following considerations:

@@

We place a cosine lobe scaled with the integrated (accumulated) energies in each of the 8 cone directions, whereas outgoing integrated and summed energies are weighted by the linear combination weights (factors) in overlapping (green) areas to split up the contributions accordingly. The contribution areas are depicted in Fig. \ref{scheme}, the cosine lobes can be observed in Fig. \ref{iter} respectively. This manner of propagation is inspired by the propagation scheme of Dachsbacher et al. \cite{dachsbacher} (see the paper on how to ``project the flux into a point light''). Basically, we sum up the weighted intensities corresponding to the neighbor respectively and scale an oriented cosine lobe with the total flux $\Phi_t$ accumulated to transmit all of the parsed intensity. The implementation is done in a dual buffer approach which pushes the energies back and forth (buffer A to buffer B) until convergence, which happens when the energy is spreaded throughout the grid and enters a stationary state, characterized by equal out (at grid borders) to in (at light source positions) energy-flow with no more nettings going on inside the domain. We measure the convergence by setting up an overall distribution error w.r.t to the previous iteration. Note that even though we propagate the energy directly through the diagonal cones (which forms a square initially), the 8 degrees of freedom introduce a circular (spherical) wave front after few iterations (for greater resolutions) as depicted in Fig. \ref{iter}. This occurs, since the diagonals are more strongly attenuated by default and there is a redistribution of intensities in the long term resulting in a circular torus-shaped distribution, as also observed in Fig. \ref{iter}. Also note, that this approach satisfies energy conservation and propagation attenuation principles and applies to the principles of light propagation in vacuum (by default). The verification of these principal requirements and specifications can be observed in the evaluation in chap. \ref{chap:eval}.

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=1.0\textwidth]
    {img/steps_gallery_format_arrow_alt.png}
  }
  \caption{Impulse response for isotropic stimulus (few iterations: $t$)}
  \label{iter}
\end{figure}\vskip 3pt
Fig \ref{iter} depicts the named iteration inidices 0,1,2,3,7,11 for the propagation scheme for a circular pulse stimulus which gets turned off after 1 iteration to demonstrate raw energy spreading in the field (impulse response of the algorithm). In step 1 one can observe the placing of cosine lobes scaled with the respective accumulated and weighted energy for the summed area in each of the 8 neighbor directions. In the subsequent steps, remark the evolution from an 8-neighborhood in step 1 to a circular shaped torus.
